User-customized hashtag/tag prediction on Twitter and StackOverflow:\protect \\ Comparing vector-based and ACT-R Bayesian models of tag retrieval

### 2013-10-06 writeups/proposal/proposal.tex
foo
Former-commit-id: 35c0f6e625a76a0b78e2092b5e39f7ebba16dd29 [formerly ff85ea3711fd6879f7dfd3c7a9d923c4e3fc58d3]
Former-commit-id: f7fb7da7afc3a81998729bad1bb9d5317fa40b39

Co-occurrence-based modeling has been shown as a potentially useful approach when predicting Twitter hashtag use \parencite{Efron2010}.
Several memory retrieval models are based on this broad methodology, where a count is maintained of the times each contextual element (such as a word in a post) co-occur with a tag.
Two of the current state-of-the-art memory models are based on a co-occurrence methodology, and I'll be comparing these two models for hashtag prediction:
ACT-R's declarative memory retrieval theory and random permutation vector-based memory systems.

@@

Alongside testing these declarative memory retrieval models on Twitter hashtag use, they will also be tested on tag use for StackOverflow posts.
StackOverflow is a question and answer site for computer programming where users ask programming-related questions and fellow members of the community provide answers.
A user with a programming-related question creates a post with their question and tags the post with a few specific programming-related tags. 
User-chosen tags on the StackOverflow site represent the primary topic of the question, such as a specific programming language, tool, software package, or framework.
Some example tags for StackOverflow are \emph{PHP}, \emph{Arrays}, \emph{MVC}, \emph{C\#}, or \emph{Common Lisp}.

@@

ACT-R's declarative memory and vector-based memory systems are substantially different in their implementation.
Nonetheless, both can successfully model classic behavioral patterns found in word pairing experiments (e.g., the ``fan effect'' \parencite{Rutledge2008}).
I will extend the work done by \textcite{Rutledge2008} by formally comparing these two models on large-scale hashtag prediction tasks.
We'll also replace the vector-based model used by \textcite{Rutledge2008} with an improved and simplified random-permutation model \parencite{Sahlgren2008}.
The primary objective is to compare the random-permutation vector-based model to ACT-R's declarative memory theory.

@@

A user's prior hashtag use will be an essential model component for domains such as Twitter, where the number of possible hashtags is practically infinite.
In these domains, the model can utilize a user's prior hashtag use as a way to prune the infinite space of possible hashtags, and generate a much smaller and more manageable set for prediction.
However, not only should the model properly take into account a user's prior hashtag use, but it should also generate new hashtag predictions when a user is most likely choosing a hashtag they have never used before.
Calibrating the models to properly balance this exploitation of prior hashtag use versus exploration of new hashtags will be a primary research focus.
The end goal is to have a model that can properly balance the user's prior tag with the contextual cues in the post in order to generate an accurate prediction.

@@

Recommendation systems based on the content of user-created posts have been successful in recommending people on Twitter for a user to follow.
\textcite{Hannon2010} built a Twitter follower recommendation system by creating a profile of each user based on the words they used in previous posts.
The recommendation system would then suggest other Twitter users to follow that had similar profiles to a particular user.
The profiles were created by generating a word frequency matrix of words x (user + followers + followees), and then providing that matrix as input into the Lucine text search engine framework.
When the system needed to recommend a set of individuals for a user to follow, it would take all of the words used in posts from that user to construct a search query to be provided to Lucine.
That search query would return the most likely users that share similar words in posts within their Twitter network with the words provided in the query.
Using this simple technique based on a standard text search engine allowed the recommendation system to accurately predict the people that a particular user will follow at 25\% accuracy.

@@

\textcite{Kuo2011} was the first to model tag use for StackOverflow posts.
His set of models were originally designed for next-word prediction for large text document collections.
In order to structure the models to work for StackOverflow tag prediction, the models would request the most-likely next word after a user created a post, and restrict the space of possible words to only tags.
He tested the set of next-word prediction models to generate predictions for the most likely tags that a user would choose when asking a question on the site.
Interestingly, the Bayesian co-occurrence model outperformed the more complicated and computationally expensive Singular Value Decomposition and K-nearest-neighbors models.
This model can accurately predict a user-chosen tag at 47\% accuracy.

@@

Google is apparently highly confident in the accuracy of their system.
On the Latex site, the user interface provides subdued suggestions to the user, but requires the user to explicitly choose each tag he/she wants to associated with the post.
For Google+, the user interface automatically provides tags for a user's post, and then requires the user to explicitly change those tags if he/she wants to use a different set.
If tag accuracy is high enough, the Google+ technique is certainly favorable since the user does not have to spend time figuring out which tags best represent this post.
However, if the model makes even just a few errors, users may become annoyed and frustrated with the recommendation system since they will have to explicitly correct it every time they generate a post.

@@

For some domains it is reasonable to assume that the base-level activations of each chunk within a particular time window of interest are stable, and do not change within that window.
Programming-language popularity over the past few years is a reasonable example of this.
Although the popularity of various programming languages has certainly changed slightly over the past few years (e.g., \emph{closure}'s growth), it is not the case that the changes have been drastic.
\emph{C\#}, \emph{Python}, and \emph{Java} are still a few of the most popular languages, \emph{Common Lisp} has not gained or lost much ground in popularity over the past few years.

@@

Strength of association ($S_{ji}$) reflects the amount of log odds adjustment to the activation of a chunk, given the current context (i.e., external environment and internal state).
Context for the StackOverflow domain for example, can be represented as each word in the title and body of a post.
Context for the Twitter domain would be the words in a Tweet.
Association strength between a chunk in memory and a single contextual element can be computed directly by calculating its context-adjusted odds ratio:
The likelihood a chunk occurred with the current context ($p(i|j)$) over the likelihood that the chunk occurred in any of the other contexts ($p(i|\overline{j})$).
For large datasets, the likelihood that a chunk occurs in any particular context reaches near-zero values ($p(j) \Rightarrow 0$).
So it can be assumed that the likelihood that a chunk occurs in any context but one ($p(i|\overline{j})$) is equivalent to the likelihood that a chunk occurs in any context ($p(i)$).
This assumption has been both mentioned and used when deriving the $S_{ji}$ equation \parencite{Anderson1989}, as well as in recent research working with large-scale datasets \parencite{Stanley2013}.

@@

The SNIF-ACT framework is also based on the large-scale version of $S_{ji}$, and scales well when working with a large-scale search query database for internet search \parencites{Fu2007,Pirolli2003}.
The $S_{ji}$ values were based on a 118GB local database of 10 million pages of the Stanford WebBase Project \parencite{Farahat2004}.
So this implementation did not fit completely in RAM, but a spinning-disk database implementation still scaled well enough for reasonably fast retrieval times for $S_{ji}$ values.

@@

\begin{table}[!ht]
  \caption{The StackOverflow tag prediction model}
  \label{tabModACTRModel}
  {\tabulinesep=1.2mm
    \begin{tabu}{ll}
      \hline
      Common Name &  Equation \\
      \hline
      Activation & 		$A_{i} = B_{i} + \sum_{j\in T}^{ } W_{j} S_{ji} + \sum_{j\in B}^{ } W_{j} S_{ji} - O$ \\
      Base Level & 		$B_{i} = log \frac{p_{i}}{1-p_{i}}$ \\
      Strength Assoc. &		$S_{ji} = log \frac{p(i|j)}{p(i))} = log \frac{NN_{ji}}{N_{Row}(j)N_{Col}(i)}$ \\
      Co-occurrence &		$N_{ji} = \sum_{}^{}{observed_{ji}}$ \\
      Observations &		$N = \sum_{}^{}{\sum_{}^{}{N_{ji}}}$ \\
      Attentional Weight& 	$W_{j} = W \frac{E_{j}} {\sum_{}^{} {E_{j}}} $ \\
      Scaled Entropy & 		$E_{j} = 1-\frac{H_{j}}{H_{max}}$ \\
      Entropy & 		$H_{j} = -\sum_{i=1}^{N}p(i|j)log\left (  p(i|j) \right )$ \\
      Offset & 			$O = \frac{1}{5}\sum_{i\in top 5}^{ } A_{i}$ \\
      \hline
    \end{tabu}
  }
\end{table}

A formal description of the modified ACT-R Bayesian model from \textcite{Stanley2013} is included in Table \ref{tabModACTRModel}.
The activation of each tag ($A_{i}$) is a function of four components:
That tag's base level activation ($B_{i}$), contextual activation for the words in the title, activation for words in the body, and an offset term ($O$).
This model has been modified from the standard ACT-R model (Table \ref{tabACT-RModel}) in three ways:
Two separate contextual components were used (words in title and body of post), an offset term was added, and an entropy measure was used to compute the attentional weight ($W_{j}$).

@@

Usually information is lost when the data are restructured into a smaller set of latent dimensions.
However it was shown that this process actually improves productiveness and reduces noise by allowing the frequency count of highly similar concepts (words) to be pooled together and represented by a single dimension.
They tested the LSA model on the Test of English as a Foreign Language (TOEFL) task of choosing the best of four possible synonyms to a target word.
Model results were on par with a large sample of applicants to american universities from non-English speaking countries (64.4\% compared to 64.5\%).
Model performance also improved when the original dimensionality of the data was reduced, but only up to a point.
Afterwards further reductions in the dimensionality of the data began to degrade performance.

@@

LSA uses the Singular Value Decomposition process to reduce the dimensionality of the word x document co-occurrence matrix.
The technique attempts to maximize the amount of variability in the data left after forcing the original dimensionality of the data to be reduced.
The best way to remove a dimension while maintaining all variability is to look for words with highly-similar co-occurrence patterns and deduce that these words represent the same concept.
For example, synonyms such as \emph{\#photograph} and \emph{\#photo} would appear interchangeably across documents, so these terms would be collapsed to a single dimension and frequency counts would be pooled.
By pooling co-occurrence counts the model is able to generate better predictions with less data.

@@

Another issue with LSA is that the SVD matrix decomposition technique is computationally very expensive.
As the size of the original word x document matrix increases, the SVD computation becomes more and more time consuming.
If the domains where LSA is being applied contained small enough datasets where SVD computation could be computed in a reasonable amount of time, then this would not be much of a concern.
However, with large-scale human-created datasets such as Twitter, Stackoverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
\textcite{Budiu2007} used the first six million pages of the Stanford WebBase corpus to compare performance between LSA and PMI on the TOEFL, and the LSA model could not be implemented with this size dataset.
Instead, they compared the measures on the smaller Touchstone Applied Science Associates (TASA) corpus created from \num{60527} samples of text from high school to early college textbooks.
This is one of the available corpora on the LSA website and is aimed to summarize the general reading up to the 1st year of college \parencite{Budiu2007}.

@@

Another issue with LSA is that adding additional observations to an already-computed reduced representation requires rerunning the SVD \textcite{Farahat2004}.
Since computing the SVD is the most computationally-expensive operation with LSA, it is difficult to see how LSA could be implemented in incremental domains where information is periodically being added.
For Twitter and StackOverflow, it might certainly be useful to update the strength of association representation when new information arrives, especially for Twitter, where hashtag use and news content change rapidly.

@@

``holographic'' memory systems \parencite{Plate1995} or vector-based memory systems represent a concept (i.e., chunk) as a vector.
The representation of the concept is distributed across all elements of the vector.
This is somewhat similar to taking a column (i.e., a tag) on a word co-occurrence matrix and saying that the distribution of co-occurrence counts across the rows in the column is the representation of that concept.
However, vector-based memory systems represent information in a much more compact way than a full word co-occurrence matrix.

@@

Retrieving the most likely chunk given context can be done in two ways: decoding and resonance \parencite{Jones2007}.
Decoding takes the memory vector for context and decodes it back into an environment vector.
The cosine between that decoded environment vector and all other environment vectors is computed and ranked, and the chunk with the environment vector with the highest cosine is retrieved.
For example, assume that a user has written the word ``zend'' in a post.
``zend'' has co-occurred with \emph{PHP}, so the unordered memory vector for ``zend'' ($c_{zend}$) contains the unordered environment vector for \emph{PHP} ($e_{PHP}$).
Since the environment vector for \emph{PHP} was added to the memory vector for ``zend'' without any encoding, no decoding is necessary.
So the cosine between the environment vector \emph{PHP} ($e_{PHP}$) and decoded memory vector for ``zend'' ($c_{zend}$) will be high, and most likely \emph{PHP} will be retrieved through decoding.

@@

One of the strengths of vector-based systems is that word order can naturally be represented alongside word co-occurrence in these distributed representations.
Representing word order is an example where both the encoding and decoding operations are applied.
As a motivating example, say that one wants to encode the first and second words immediately preceding a hashtag in a tweet.
Simply counting co-occurrences of the two immediately-preceding words to hashtags does not include the order information of the words in the representation.
With this representation, it would be impossible to tell the difference between 1-back words that immediately preceded the hashtag, and 2-back words that occurred just prior to these 1-back words.
Order information is lost if a simply co-occurrence counting technique is used.

@@

For example, suppose that a retrieval request is made for ``martin $\Phi$'' given that the model was trained on the TASA corpus.
Within this corpus, the phrase ``martin luther'' often occurred.
When this occurs, the word-order memory vector for martin ($o_{martin}$) will be updated with the convolving luther and a 1-forward location vector ($o_{martin} = o_{martin'} + \Phi * e_{luther}$).
Also, the word-order memory vector for luther ($o_{luther}$) will be updated by convolving martin and a 1-back location vector ($o_{luther} = o_{luther'} + e_{martin} * \Phi$).
The retrieval request for ``martin $\Phi$'' can be done in two ways: through decoding or resonance.

@@

\textcite{Recchia2010} found that random permutations could use smaller vectors to produce equivalent performance (better compression), and scale to larger datasets (better computational efficiency).
The vector length x performance tradeoff was tested on a simple paired associative retrieval task. 
Retrieval accuracy for the vector-based models remained high even after reducing each environment vector's length from \num{2048} to \num{1024}.
This indicates that vector-based representations using random permutations compress better than those using circular convolution.

To test computational efficiency, each model was trained on a subset of the Wikipedia corpus, and tested on the Test of English as a Foreign Language (TOEFFL) assessment (among other tasks).
The random-permutation model was trained on \num{2.33} GB corpus.
The BEAGLE model was unable to train on a corpus this large, so both models were also trained on an additional \num{35} MB subset of the Wikipedia corpus.
The random permutation model performed as well or slightly better than the BEAGLE model when trained on the smaller corpus.
The model's performance also significantly improved when trained on the much larger corpus that was computationally intractable for the BEAGLE model.
These results make a strong case to favor random-permutations over circular convolution when implementing vector-based memory representations.

@@

Information is lost with this compression in both cases: by starting with a compressed representation with random permutations or by reducing to a compressed representation with LSA.
However, the way that the compression works in these two cases is not the same.
LSA uses SVD to find words that have highly correlated tag occurrence distributions (i.e., latent concepts), and then pools those co-occurrence counts into a single dimension.
A vector-based approach like BEAGLE or random permutations essentially randomizes the mapping of the words on the reduced set of dimensions, and does not take into account word similarity when generating the mapping.
Nonetheless, random permutations have been shown to perform similar to if not better than LSA when trained on the TASA and tested on the TOEFFL \parencites{Sahlgren2008,Jones2007}.

@@

However, vector-based models and ACT-R's current Bayesian DM model are not identical, and there are open issues with both models that need to be addressed before further evaluating the benefits of one over the other.
For example, the frequency and recency of retrievals for each chunk are not currently used for vector-based models when calculating activation.
More recent tweets on a news site like Twitter should be more relevant, for example.
\textcite{Efron2011} tested this claim by evaluating a set of query likelihood models on returning relevant Twitter tweets from search queries.
They showed that a model incorporating recency information of the tweet produced more relevant results than models that did not pay attention to this information.

>>>>

Co-occurrence-based modeling has been shown as a potentially useful approach when predicting Twitter hashtag use \parencite{Efron2010}.
Several memory retrieval models are based on this broad methodology, where a count is maintained of the times each contextual element (such as a word in a post) co-occur with a tag.
Two of the current state-of-the-art memory retrieval models are based on a co-occurrence methodology, and I'll be comparing these two models for hashtag prediction:
ACT-R's declarative memory retrieval theory and random permutation vector-based memory systems.

@@

Alongside testing these memory retrieval models on Twitter hashtag use, they will also be tested on tag use for StackOverflow posts.
StackOverflow is a question and answer site for computer programming where users ask programming-related questions and fellow members of the community provide answers.
A user with a programming-related question creates a post with their question and tags the post with a few specific programming-related tags. 
User-chosen tags on the StackOverflow site represent the primary topic of the question, such as a specific programming language, tool, software package, or framework.
Some example tags for StackOverflow are \emph{PHP}, \emph{Arrays}, \emph{MVC}, \emph{C\#}, and \emph{Common Lisp}.

@@

ACT-R's declarative memory and vector-based memory systems are substantially different in their mathematical structure.
Nonetheless, both can successfully model classic behavioral patterns found in word pairing experiments (e.g., the ``fan effect'' \parencite{Rutledge2008}).
I will extend the work done by \textcite{Rutledge2008} by formally comparing these two models on large-scale hashtag prediction tasks.
I'll also replace the vector-based model used by \textcite{Rutledge2008} with an improved and simplified random-permutation model \parencite{Sahlgren2008}.
The primary objective is to compare the random-permutation vector-based model to ACT-R's declarative memory theory.

@@

A user's prior hashtag use will be an essential model component for domains such as Twitter, where the number of possible hashtags is practically infinite.
In these domains, the model can utilize a user's hashtag history as a way to prune the infinite space of possible hashtags, and generate a much smaller and more manageable set for prediction.
However, not only should the model properly take into account a user's prior hashtag use, but it should also generate new hashtag predictions when a user is most likely choosing a hashtag they have never used before.
Calibrating the models to properly balance this exploitation of prior hashtag use versus exploration of new hashtags will be a primary research focus.
The end goal is to have a model that can properly balance the user's prior tag with the contextual cues in the post in order to generate an accurate prediction.

@@

Recommendation systems based on the content of user-created posts have been successful in recommending people on Twitter for a user to follow.
\textcite{Hannon2010} built a Twitter follower recommendation system by creating a profile of each user based on the words they used in previous posts.
The recommendation system would then suggest other Twitter users to follow that had similar profiles to a particular user.
The profiles were created by generating a word frequency matrix of words x (user + followers + followees), and then providing that matrix as input into the Lucine text search engine framework.
When the system needed to recommend a set of individuals for a user to follow, it would take all of the words used in posts from that user to construct a search query to be provided to Lucine.
That search query would return the most likely users that share similar words in posts within their Twitter network with the words provided in the query.
Using this simple technique based on a standard text search engine allowed the recommendation system to accurately predict the people that a particular user follows at 25\% accuracy.

@@

\textcite{Kuo2011} was the first to model tag use for StackOverflow posts.
His set of models were originally designed for next-word prediction for large text document collections.
In order to structure the models to work for StackOverflow tag prediction, the models would request the most-likely next word after a user created a post, and restrict the space of possible words to only tags.
He tested the set of next-word prediction models to generate predictions for the most likely tags that a user would choose when asking a question on the site.
Interestingly, the Bayesian co-occurrence model outperformed the more complicated and computationally expensive Singular Value Decomposition and K-nearest-neighbors models.
This Bayesian model can accurately predict a user-chosen tag at 47\% accuracy.

@@

Google is apparently highly confident in the accuracy of their system.
On the Latex site, the user interface provides subdued suggestions to the user, but requires the user to explicitly choose each tag he/she wants to associated with the post.
For Google+, the user interface automatically tags a user's post, and then requires the user to explicitly change those tags if he/she wants to use a different set.
If tag accuracy is high enough, the Google+ technique is certainly favorable since the user does not have to spend time figuring out which tags best represent this post.
However, if the model makes even just a few errors, users may become annoyed and frustrated with the recommendation system since they will have to explicitly correct it every time they generate a post.

@@

For some domains it is reasonable to assume that the base-level activations of each chunk within a particular time window of interest are stable, and do not change within that window.
Programming-language popularity over the past few years is a reasonable example of this.
Although the popularity of various programming languages has certainly changed slightly over the past few years (e.g., Clojure's growth), it is not the case that the changes have been drastic.
C\#, Python, and Java are still a few of the most popular languages, Common Lisp has not gained or lost much ground in popularity over the past few years.

@@

Strength of association ($S_{ji}$) reflects the amount of log odds adjustment to the activation of a chunk, given the current context (i.e., external environment and internal state).
Context for the StackOverflow domain for example is represented as each word in the title and body of a post.
Context for the Twitter domain are the words in a Tweet.
Association strength between a chunk in memory and a single contextual element can be computed directly by calculating its context-adjusted odds ratio:
The likelihood a chunk occurred with the current context ($p(i|j)$) over the likelihood that the chunk occurred in any of the other contexts ($p(i|\overline{j})$).
For large datasets, the likelihood that a chunk occurs in any particular context reaches near-zero values ($p(j) \Rightarrow 0$).
So it can be assumed that the likelihood that a chunk occurs in any context but one ($p(i|\overline{j})$) is equivalent to the likelihood that a chunk occurs in any context ($p(i)$).
This assumption has been both mentioned and used when deriving the $S_{ji}$ equation \parencite{Anderson1989}, as well as in recent research working with large-scale datasets \parencite{Stanley2013}.

@@

The SNIF-ACT framework \parencites{Fu2007,Pirolli2003} is also based on the large-scale version of $S_{ji}$, and scales well when working with a large-scale search query database for internet search.
\textcite{Farahat2004} used $S_{ji}$ values based on a 118GB local database of 10 million pages of the Stanford WebBase Project.
So this implementation did not fit completely in RAM, but a spinning-disk database implementation still scaled well enough for reasonably fast retrieval times for $S_{ji}$ values.

@@

A formal description of the modified ACT-R Bayesian model from \textcite{Stanley2013} is included in Table \ref{tabModACTRModel}.

\begin{table}[!ht]
  \caption{StackOverflow tag prediction model}
  \label{tabModACTRModel}
  {\tabulinesep=1.2mm
    \begin{tabu}{ll}
      \hline
      Common Name &  Equation \\
      \hline
      Activation & 		$A_{i} = B_{i} + \sum_{j\in T}^{ } W_{j} S_{ji} + \sum_{j\in B}^{ } W_{j} S_{ji} - O$ \\
      Base Level & 		$B_{i} = log \frac{p_{i}}{1-p_{i}}$ \\
      Strength Assoc. &		$S_{ji} = log \frac{p(i|j)}{p(i))} = log \frac{NN_{ji}}{N_{Row}(j)N_{Col}(i)}$ \\
      Co-occurrence &		$N_{ji} = \sum_{}^{}{observed_{ji}}$ \\
      Observations &		$N = \sum_{}^{}{\sum_{}^{}{N_{ji}}}$ \\
      Attentional Weight& 	$W_{j} = W \frac{E_{j}} {\sum_{}^{} {E_{j}}} $ \\
      Scaled Entropy & 		$E_{j} = 1-\frac{H_{j}}{H_{max}}$ \\
      Entropy & 		$H_{j} = -\sum_{i=1}^{N}p(i|j)log\left (  p(i|j) \right )$ \\
      Offset & 			$O = \frac{1}{5}\sum_{i\in top 5}^{ } A_{i}$ \\
      \hline
    \end{tabu}
  }
\end{table}

The activation of each tag ($A_{i}$) is a function of four components:
That tag's base level activation ($B_{i}$), contextual activation for the words in the title, activation for words in the body, and an offset term ($O$).
This model has been modified from the standard ACT-R model (Table \ref{tabACT-RModel}) in three ways:
Two separate contextual components were used (words in title and body of post), an offset term was added, and an entropy measure was used to compute the attentional weight ($W_{j}$).

@@

Usually information is lost when the data are restructured into a smaller set of latent dimensions.
However it was shown that this process actually improves productiveness and reduces noise by allowing the frequency count of highly similar concepts (words) to be pooled together and represented by a single dimension.
\textcite{Landauer1997} tested the LSA model on the Test of English as a Foreign Language (TOEFL) task of choosing the best of four possible synonyms to a target word.
Model results were on par with a large sample of applicants to american universities from non-English speaking countries (64.4\% compared to 64.5\%).
Model performance also improved when the original dimensionality of the data was reduced, but only up to a point.
Afterwards further reductions in the dimensionality of the data began to degrade performance.

@@

LSA uses the Singular Value Decomposition process to reduce the dimensionality of the word x document co-occurrence matrix.
The technique attempts to maximize the amount of variability in the data left after forcing the original dimensionality of the data to be reduced.
The best way to remove a dimension while maintaining the most variability is to look for words with highly-similar co-occurrence patterns and deduce that these words represent the same concept.
For example, synonyms such as \emph{\#photograph} and \emph{\#photo} would appear interchangeably across documents, so these terms would be collapsed to a single dimension and frequency counts would be pooled.
By pooling co-occurrence counts the model is able to generate better predictions with less data.

@@

Another issue with LSA is that the SVD matrix decomposition technique is computationally very expensive.
As the size of the original word x document matrix increases, the SVD computation becomes more and more time consuming.
If the domains where LSA is being applied contained small enough datasets where SVD computation could be computed in a reasonable amount of time, then this would not be much of a concern.
However, with large-scale human-created datasets such as Twitter, Stackoverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
\textcite{Budiu2007} used the first six million pages of the Stanford WebBase corpus to compare performance between LSA and PMI on the TOEFL, and the LSA model could not be implemented with this size training dataset.
Instead, they compared the measures on the TOEFL by training on the smaller Touchstone Applied Science Associates (TASA) corpus created from \num{60527} samples of text from high school to early college textbooks.
This is one of the available corpora on the LSA website and is aimed to summarize the general reading up to the 1st year of college \parencite{Budiu2007}.

@@

Another issue with LSA is that adding additional observations to an already-computed reduced representation requires rerunning the SVD \parencite{Farahat2004}.
Since computing the SVD is the most computationally-expensive operation with LSA, it is difficult to see how LSA could be implemented in incremental domains where information is periodically being added.
For Twitter and StackOverflow, it might certainly be useful to update the strength of association representation when new information arrives, especially for Twitter, where hashtag use and news content change rapidly.

@@

``Holographic'' memory systems \parencite{Plate1995} or vector-based memory systems represent a concept (i.e., chunk) as a vector.
The representation of the concept is distributed across all elements of the vector.
This is somewhat similar to taking a column (i.e., a tag) on a word co-occurrence matrix and viewing the distribution of co-occurrence counts across the rows in the column as the representation of that concept.
However, vector-based memory systems represent information in a much more compact way than a full word co-occurrence matrix.

@@

Retrieving the most likely chunk given context can be done in two ways: decoding and resonance \parencite{Jones2007}.
Decoding takes the memory vector for context and decodes it back into an environment vector.
The cosine between that decoded environment vector and all other environment vectors is computed and ranked, and the chunk with the environment vector with the highest cosine is retrieved.
For example, assume that a user has written the word ``zend'' in a post.
``zend'' has co-occurred with \emph{PHP}, so the unordered memory vector for ``zend'' ($c_{zend}$) contains the unordered environment vector for \emph{PHP} ($e_{\mathit{PHP}}$).
Since the environment vector for \emph{PHP} was added to the memory vector for ``zend'' without any encoding, no decoding is necessary.
So the cosine between the environment vector \emph{PHP} ($e_{\mathit{PHP}}$) and decoded memory vector for ``zend'' ($c_{zend}$) will be high, and most likely \emph{PHP} will be retrieved through decoding.

@@

One of the strengths of vector-based systems is that word order can naturally be represented alongside word co-occurrence in these distributed representations.
Representing word order is an example where both the encoding and decoding operations are applied.
As a motivating example, say that one wants to encode the first and second words immediately preceding a hashtag in a tweet.
Simply counting co-occurrences of the two immediately-preceding words and hashtags does not include the order information of the words in the representation.
With this representation, it would be impossible to tell the difference between 1-back words that immediately preceded the hashtag, and 2-back words that occurred just prior to these 1-back words.
Order information is lost if a simply co-occurrence counting technique is used.

@@

For example, suppose that a retrieval request is made for ``martin $\Phi$'' given that the model was trained on the TASA corpus.
Within this corpus, the phrase ``martin luther'' often occurred.
When this occurs, the word-order memory vector for martin ($o_{martin}$) will be updated with the convolving luther and a 1-back location vector ($o_{martin} = o_{martin'} + \Phi * e_{luther}$).
Also, the word-order memory vector for luther ($o_{luther}$) will be updated by convolving martin and a 1-forward location vector ($o_{luther} = o_{luther'} + e_{martin} * \Phi$).
The retrieval request for ``martin $\Phi$'' can be done in two ways: through decoding or resonance.

@@

\textcite{Recchia2010} found that random permutations could use smaller vectors to produce equivalent performance (better compression), and scale to larger datasets (better computational efficiency).
The vector length x performance tradeoff was tested on a simple paired associative retrieval task. 
Retrieval accuracy for the random-permutation vector-based models remained high even after reducing each environment vector's length from \num{2048} to \num{1024}.
This indicates that vector-based representations using random permutations compress better than those using circular convolution.

To test computational efficiency, each model was trained on a subset of the Wikipedia corpus, and tested on the Test of English as a Foreign Language (TOEFL) assessment (among other tasks).
The random-permutation model was trained on \num{2.33} GB corpus.
The BEAGLE model was unable to train on a corpus this large, so both models were also trained on an additional \num{35} MB subset of the Wikipedia corpus.
The random permutation model performed as well or slightly better than the BEAGLE model when trained on the smaller corpus.
The model's performance also significantly improved when trained on the much larger corpus that was computationally intractable for the BEAGLE model.
These results make a strong case to favor random-permutations over circular convolution when implementing vector-based memory representations.

@@

Information is lost with this compression in both cases: by starting with a compressed representation with random permutations or by reducing to a compressed representation with LSA.
However, the way that the compression works in these two cases is not the same.
LSA uses SVD to find words that have highly correlated tag occurrence distributions (i.e., latent concepts), and then pools those co-occurrence counts into a single dimension.
A vector-based approach like BEAGLE or random permutations essentially randomizes the mapping of the words on the reduced set of dimensions, and does not take into account word similarity when generating the mapping.
Nonetheless, random permutations have been shown to perform similar to if not better than LSA when trained on the TASA and tested on the TOEFL \parencites{Sahlgren2008,Jones2007}.

@@

However, vector-based models and ACT-R's current Bayesian DM model are not identical, and these differences need to be addressed in order to better evaluate the benefits of one over the other.
For example, the frequency and recency of retrievals for each chunk are not currently used for vector-based models when calculating activation.
More recent tweets on a news site like Twitter should be more relevant, for example.
\textcite{Efron2011} tested this claim by evaluating a set of query likelihood models on returning relevant Twitter tweets from search queries.
They showed that a model incorporating recency information of the tweet produced more relevant results than models that did not pay attention to this information.
